---
title: "css_homework 2B"
output:
  html_document:
    df_print: paged
---

# Case 1: Peru

```{r}
# Load required libraries
library(igraph)
library(leiden) 
library(tidygraph) 
library(scales)  
library(ggraph)

```

```{r}
peru_url <- "https://github.com/DACSS-CSSmethods/HW2_repo/raw/refs/heads/main/case1_network.graphml"
peru_graph <- read_graph(peru_url, format = "graphml")
```

```{r}
# Convert to a tidygraph object
peru_tbl <- as_tbl_graph(peru_graph)
```

```{r}
# Add Louvain and Leiden groupings as node attributes
peru_tbl <- peru_tbl %>%
  mutate(louvain = as.factor(group_louvain()),
         leiden = as.factor(group_leiden()))

```

```{r}
# Visualize Louvain communities
ggraph(peru_tbl, layout = "fr") +
  geom_edge_link(alpha = 0.3) +
  geom_node_point(aes(color = louvain), size = 4) +
  theme_void() +
  ggtitle("Louvain Community Detection")

```

```{r}
# Visualize Leiden communities
ggraph(peru_tbl, layout = "fr") +
  geom_edge_link(alpha = 0.3) +
  geom_node_point(aes(color = leiden), size = 4) +
  theme_void() +
  ggtitle("Leiden Community Detection")

```

```{r}
# Number of communities
length(unique(peru_tbl$louvain))
length(unique(peru_tbl$leiden))
```

```{r}
# Modularity for Louvain
modularity(cluster_louvain(peru_graph))

# Modularity for Leidan
leiden_membership <- peru_tbl %>% 
  activate(nodes) %>% 
  pull(leiden) %>% 
  as.integer()

modularity(peru_graph, membership = leiden_membership)
```

```{r}
# Run Louvain and Leiden on peru_graph
louvain_obj <- cluster_louvain(peru_graph)
leiden_obj <- cluster_leiden(peru_graph)

# Assign community membership back to the graph
V(peru_graph)$louvain <- membership(louvain_obj)
V(peru_graph)$leiden <- membership(leiden_obj)

# Louvain plot
plot(peru_graph,
     vertex.color = V(peru_graph)$louvain,
     main = "Louvain Community Detection",
     vertex.label = NA, vertex.size = 8, edge.arrow.size = 0.5)

# Leiden plot
plot(peru_graph,
     vertex.color = V(peru_graph)$leiden,
     main = "Leiden Community Detection",
     vertex.label = NA, vertex.size = 8, edge.arrow.size = 0.5)

```

## Results

Louvain detected 4 clear communities with a strong modularity score of 0.29, while Leiden over-partitioned the network into 31 groups, resulting in a negative modularity score. Based on these results, Louvain is the more appropriate choice for this network.

# Case 2: Seattle

```{r}
seattle_url <- "https://github.com/DACSS-CSSmethods/HW2_repo/raw/refs/heads/main/case2_network.graphml"
seattle_graph <- read_graph(seattle_url, format = "graphml")
```

```{r}
# Louvain
#louvain_seattle <- cluster_louvain(as.undirected(seattle_graph))
#length(louvain_seattle)
#modularity(louvain_seattle)

# Leiden
#leiden_seattle <- cluster_leiden(as.undirected(seattle_graph))
#length(leiden_seattle)
#modularity(seattle_graph, membership = membership(leiden_seattle))

```

Neither Louvain or Leiden works because they cannot be applied to directed graphs. The code above results in the following error message:

```         
Error in cluster_leiden(seattle_graph) : 
  At vendor/cigraph/src/community/leiden.c:1000 : Leiden algorithm is only implemented for undirected graphs. Invalid value""
```

```{r}
# Infomap
infomap_seattle <- cluster_infomap(seattle_graph)
length(infomap_seattle)
modularity(infomap_seattle)

# Label Propagation
label_seattle <- cluster_label_prop(seattle_graph)
length(label_seattle)
modularity(label_seattle)

```

```{r}
# Normalize followers for node size
follower_scaled <- scales::rescale(V(seattle_graph)$followers, to = c(5, 15))

# Plot the graph sized by followers
plot(seattle_graph,
     vertex.size = follower_scaled,
     vertex.color = "skyblue",
     vertex.label = NA,
     edge.arrow.size = 0.3,
     main = "Seattle Network – Node Size = # of Followers")

```

## Results

The Infomap and Label Propagation algorithms detected only a single community with a modularity score of 0. This indicates that no meaningful community structure was identified. A visualization of the network suggests it may be too densely connected or structurally uniform for these algorithms to detect distinct clusters.

# Case 3: FIFA

```{r}
fifa_url <- "https://github.com/DACSS-CSSmethods/HW2_repo/raw/refs/heads/main/case3_network.graphml"
fifa_graph <- read_graph(fifa_url, format = "graphml")
```

```{r}
# Girvan–Newman
gn_fifa <- cluster_edge_betweenness(fifa_graph)
length(gn_fifa)
modularity(gn_fifa)

```

```{r}
# Leiden
leiden_fifa <- cluster_leiden(fifa_graph)
length(leiden_fifa)
modularity(fifa_graph, membership = membership(leiden_fifa))
```

```{r}
# Assign community memberships to the graph
V(fifa_graph)$girvan <- membership(gn_fifa)
V(fifa_graph)$leiden <- membership(leiden_fifa)

# Plot Girvan–Newman
plot(fifa_graph,
     vertex.color = V(fifa_graph)$girvan,
     vertex.label = NA,
     vertex.size = 6,
     main = "Girvan–Newman Community Detection")

# Plot Leiden
plot(fifa_graph,
     vertex.color = V(fifa_graph)$leiden,
     vertex.label = NA,
     vertex.size = 6,
     main = "Leiden Community Detection")

```

## Results

Girvan–Newman detected 25 meaningful communities with a modularity score of 0.35, while Leiden fragmented the network into 222 small groups with very low modularity (0.03). Based on these results, Girvan–Newman is the better choice for identifying communities in this network.
